Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-12-04 doc/om-paper/paper.tex
fixed text in synch observer section.

\usepackage{graphicx}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

@@

In this paper, we develop a formal framework to address these concerns. We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems. We describe a generic formal framework for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples. The framework is suitable for infinite-state model-checking.

@@

@@

\subsubsection{A Synchronous Kibitzer}\label{sec:kibitzer}
The typical approach to modeling faults is to add new state variables to each process representing its fault state. Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process. As a simple example, we might define a node that sends a good message if it is non-faulty and a bad message otherwise. In pseudo-code using guarded commands, its definition might look like the following:

\begin{verbatim}
node:
  health: Fault\_Type;

  faulty(health)      --> send(bad\_msg);
  non\_faulty(health) --> send(good\_msg);
\end{verbatim}

@@

\noindent
But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment. Doing so divorces a node specification from its implementation.

A second difficulty with model-checking fault-tolerant systems in general is that modeling faults requires adding state and nondeterminism. The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size. Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby:

@@

To address these issues, we introduce a synchronous observer that injects faults that we call a \emph{synchronous kibitzer}. The kibitzer decomposes the state and transitions associated with the fault model from the system. For the sake of concreteness, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{hybrid}. This fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults. It  applies to broadcast systems in which a process is expected to broadcast the same value to multiple receivers. A \emph{Byzantine fault} is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value). A \emph{symmetric fault} is one in which a process may broadcast the same, but incorrect, value to other processes, and a \emph{manifest fault} is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.

@@

Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from processes to faults.

@@

\lee{msg in SAL is recv here}

@@

The upshot is that the fault model and fault-injection is completely modular to the specification of the protocol and node behavior.

>>>>

\usepackage{graphicx}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

@@

In this paper, we develop a formal framework to address these concerns. We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems. We describe a generic formal framework for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples. The framework is suitable for infinite-state model-checking supported by satisfiability modulo theories (SMT).

@@

In specifying a system, a \emph{clock} module is synchronously composed with the system. The clock module updates time to the next calendar event (and stutters otherwise.

@@

\subsection{A Synchronous Kibitzer}\label{sec:kibitzer}
The typical approach to modeling faults is to add new state variables to each process representing its fault state. Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process. As a simple example, we might define a node that sends a good message if it is non-faulty and a bad message otherwise. In pseudo-code using guarded commands, its definition might look like the following:

\begin{verbatim}
node:
  health: Fault_Type;
  faulty(health)      --> send(bad_msg);
  non_faulty(health) --> send(good_msg);
\end{verbatim}

@@

\noindent
But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment. Generally, nodes do not contain state variables assigned to their faults, or use their fault-status to determine their behavior!\footnote{There are exceptions. Benign faults may be detected by a node itself (e.g., in a built-in-test), an distributed diagnosis can let a node discover that others believe it is faulty~\cite{}.} The upshot is that combining faults and node state divorces the specification from its implementation.

A second difficulty with model-checking fault-tolerant systems in general is that modeling faults requires adding state and nondeterminism. The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size. Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby in describing the number of data values that a model must include to model the full range of Byzantine faults (defined later in this section):

@@

The second problem is the most straightforward to solve. In infinite-state model-checking, we can use either the integers or the reals as the datatype for values. Fault-tolerant voting schemes, such as a majority vote or mid-value selection (see Section~\ref{sec:byz}), require only equality, or a total order, respectively, to be defined for the data.

The solution to the first problem is more involved. Our solution is to introduce what we call a \emph{synchronous kibitzer} that symbolically injects faults into the model. The kibitzer decomposes the state and transitions associated with the fault model from the system itself. For the sake of concreteness in describing the synchronous kibitzer, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{hybrid}. This fault model distinguishes Byzantine, symmetric, and manifest faults. It  applies to broadcast systems in which a process is expected to broadcast the same value to multiple receivers. A \emph{Byzantine} (or \emph{arbitrary} fault is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value). A \emph{symmetric} fault is one in which a process may broadcast the same, but incorrect, value to other processes. Finally, a \emph{manifest} (or \emph{benign} fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the valuep arrives outside of a predetermined window.

@@

Define a set of fault types $$\textnormal{Faults} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{Faults}$$ range over possible mappings from processes to faults.

@@

@@

\paragraph{Transient faults.} If the $faults$ mapping is a constant, then faults are permanently but nondeterministically assigned to nodes. However, we can easily model \emph{transient faults}~\cite{} in which nodes are faulty temporarily by making $faults$ a state variable that is updated nondeterministically. Whether we model permanent or transient faults, a \emph{maximum fault assumption} (MFA) describes the maximum number of faults permitted in the system. The $faults$ mapping can be nondeterministically updated while satisfying the MFA using a constraint such as $faults \in \{f \, | \, mfa(f) \}$, where the MFA is defined by $mfa()$.
